Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2020-08-19 write-up/fda-recurrent.tex
Getting closer: need to get sims finished and the application done

During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital were enrolled in an observational study.  Study data were graciously provided by Evan Kleiman and his study team ({\tt https://kleimanlab.org}). Each study participant wore an Empatica E4~\citep{empaticae3}, a medical-grade wearable device that offers real-time physiological data.  On each user-day, participants were asked to self-identify moments of suicidal distress.  At these times, the participant pressed a button on the Empatica E4 device and the moment was recorded.  One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress.

The
distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement. The scientific hypothesis to be tested is whether a spike in EDA and/or gACC

To avoid, we focus on the study window 9AM to 8PM each study day.

At the end of the study period, the base rate of button pressing  rate was 292/488, or approximately 60\%.

\begin{figure}[!th]
\caption{Button press distribution across study days (include number of people).}
\end{figure}

The focus here is on two EDA and the  prothrombin index, a composite
blood coagulation index
related to liver function, measured initially at three-month intervals
and subsequently at roughly twelve-month intervals.
The individual EDA and AI trajectories are highly variable, both in forward and when considering reverse time from event timepoints, which tends to obscure patterns and trends.  In Figure~XX, the mean trajectory is plotted against time from recruitment for two patient groups placebo/prednisone and censored/not censored. Naturally, only those patients who are still alive are included in the average
for that time.
Figure XX hows the same plots in reverse alignment.
While there are certain similarities in the two plots,
the differences in temporal trends are rather striking.
In particular, prothrombin levels in the six months prior to censoring are fairly stable,
which is in marked contrast with levels in the six months prior to failure,
as seen in the lower pair of curves.

\begin{figure}[!th]
\caption{EDA and ACC plots for average at non-event and event times.}
\end{figure}

Inspection of the graphs for uncensored patients in the right panel of Figure XX suggests beginning with the simplest revival model in which the sequences
for distinct patients are independent Gaussian with moments {\bf comparison over $\Delta$ and using imputation}

@@

>>>>

During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital were enrolled in an observational study.  Study data were graciously provided by Evan Kleiman and his study team ({\tt https://kleimanlab.org}). Each study participant wore an Empatica E4~\citep{empaticae3}, a medical-grade wearable device that offers real-time physiological data.  On each user-day, participants were asked to self-identify moments of suicidal distress.  At these times, the participant was asked to press a button on the Empatica E4 device.  The timestamp of the button press was recorded.  One of the primary study goals was to assess the association between sensor-based physiological measurements and self-identified moments of suicidal distress. In particular, the scientific question is whether there are early indicators of escalating distress by monitoring physiological correlates.

A key concern is whether all moments of suicidal distress are recorded.  To ensure this, clinical staff interviewed participants in the evening who were then asked to review their button press activity.  Any events that were identified as incorrect button press activity were removed.  An additional concern is that events can only occur when the individual is at-risk, i.e., (A) the individual is wearing the Empatica E4 and (B) is currently awake.
Due to (B), we define each study day to begin at 9AM and end at 8PM. Due to (A), we use the Empatica E4 temperature sensor and filter out all records during which temperature data are below $30^{\circ}$ Celsius. At the end of the study period, the average number of button presses per day was XX with a standard deviation of XX.  Figure~\ref{fig:buttons} presents the distribution of button presses as a function of day in study.

\begin{figure}[!th]
\caption{Button press distribution across study days (include number of people).}
\label{fig:buttons}
\end{figure}

The focus here is on two physiological processes -- (1) electrodermal activity (EDA), a measure of skin conductance measured at 4Hz, and  (2) \emph{gACC}, a coordinate-free feature built from accelerometer data collected at 32Hz that measures physical movement.  The EDA measures are cleaned to  Due to high between study-day variation in EDA which may be due to external factors, we analyze EDA standardized per study-day.

The individual EDA and gACC trajectories are highly variable which tends to obscure patterns and trends.  In Figure~XX, the mean trajectories of EDA and gACC are plotted in reverse time from button press timestamps and randomly sampled non-event timestamps. While there are certain similarities in the two plots, the differences in temporal trends are rather striking. In particular, we see sharp increases in EDA in the XX minutes prior to button presses,
which is in marked contrast with levels in the six months prior to failure,
as seen in the lower pair of curves.

\begin{figure}[!th]
\caption{EDA and ACC plots for average at non-event and event times.}
\label{fig:means}
\end{figure}

Inspection of the graphs suggests setting $\Delta = XX$ for EDA and $XX$ for gACC and  Gaussian with moments {\bf comparison over $\Delta$ and using imputation} where the baseline rate is assumed to depend on

@@

1. Start simulation for appendix (delta wrong, beta2)
2. Button Press Figure
3. Compute wear minutes
4. Standardize EDA
5. Subsample EDA and gACC only at wear times between 9am and 8pm
6. EDA/ACC figures
7. Estimate with complete data
8. Estimate with bootMI; make 1 plot with CIs for both compl and bootMI
9. Estimate with heterogeneity
10. Present
